### 2020-03-18 draft/response.txt
response draft

Response to review 1:

- "it would improve the paper to give more of an intuitive understanding"

We agree and will try to find space for this. A possible intuitive explanation
is that in KKA the universe U was empty, and all substitutions were "neutral",
i.e. semantic subsitutions were functors which permute, duplicate or forget
components of algebras, or freely reinterpret components, and all such functors
strictly preserve limits. In the current paper, U is not empty, it's closed
under identity and infinitary function types, which don't just simply reshuffle
structure and hence preserve limits only weakly.

- "As stated this is only a "logical equivalence,""

We will note this. Isomorphism is probably not feasible, while impredicative
Church codings work to some extent (see "Impredicative encodings of HITs" by
Awodey et al.), they don't support large elimination, and we need interpretation of
signatures in large models.

Response to review 2:

- "it would be good to have a story how to work with QIITs in practice"

We didn't elaborate QIIT usage and applications because of space constraints.
However, we don't think "escaping from quotients" is an issue in practice. We
can print quotients as a propositionally truncated strings, which work just as
well as printing untruncated strings. Likewise with first digits of real
numbers. We agree though that working with quotients requires at least some
deliberation and rephrasing.

- "El (Π^{inf} A b) = Π^{ext} A (El ∘ b)"

This equation plausibly holds in the M model, we have not included/checked it
because it is not really important with respect to practical usage of signatures.

- (A : Ty_Γ₀ Γ) → Tm_Γ₀ Γ A

We'll try to add more illustration. Here probably the best would be to note the
following:

  Con_Γ        type of Γ-algebras
  Sub_Γ Γ Δ    type of homomorphisms from Γ to Δ
  Ty_Γ Γ       type of displayed algebras over Γ
  Tm_Γ Γ A     type of dependent morphisms from Γ to A

If Γ₀ is the CwF of Nat-algebras, then Γ : Con_Γ₀ is a particular Nat-algebra
(N, z, s), A has type (Nᴾ : N → Set) × (zᴾ : Nᴾ z) × (sᴾ : (n : N) → Nᴾ n → Nᵖ
(s n)) and Tm_Γ₀ Γ A is a type which is equal to (Nˢ : (n : N) → Nᴾ n) × (Nˢ z =
zᴾ) × ((n : N) → Nˢ (s n) = sᴾ n (Nˢ n)). It could be illustrative as well to
see how "initiality" looks like in the same setting:

  Initial Γ := (Δ : Con_Γ) → isContr (Sub_Γ Γ Δ)

"Inductive" is a dependently typed analogue of "Initial" in the sense that there
is a dependent morphism (Tm Γ A) from an inductive object Γ to any A dependent
object over it.

Response to review 3:

- "Bootstrapping how exactly?"

The benefit of bootstrapping is that the only ToS-related thing which we have to
manually invent is the notion of ToS model in def 3.1, and notions of
ToS-induction, ToS-morphism, etc. are computed by the semantics of ToSSig. The
induction principle for ToS is rather huge (it says that every displayed ToS
algebra has a section), and it was a bit unsatisfactory in the previous
"Constructing QIITs" paper that it was ad-hoc assumed/invented.

- "the signature ToSSig helps you
   with the notion of ToS-induction, but what is the concrete link with
   the assumed initial model syn in Section 6?"

The concrete link is that we need a notion of initiality before we can assume an
*initial* ToS model. The M interpretation of ToSSig computes (among other
things) this. ToSSig is defined in a way so that the Con part of its M
interpretation yields exactly the notion of ToS algebra from def 3.1, so that we
get an flCwF where objects are ToS algebras and morphisms are strict ToS
morphisms.

- "Why finite limit CwFs?"

The equivalence of induction and initiality (which we reference from
"Constructing QIITs") requires some (not all) components of CwF + identity +
constant families. We add all CwF+Id+K+Sigma components so that we get the
structure which is (bi)equivalent to finitely complete categories. There are also
many potentially useful/interesting constructions on such CwFs in the works of
Dybjer, Clairambault & Castellan, so we think that it's worth to support this
structure. We'll try to make this motivation clearer

- "Making sense of Section 6"
- "but the flCwF structure seems to be missing"

There is no fl-structure involved here, we define three (displayed) models of
ToS, the set model, the logical predicate model and the section model. We agree
that it is hard to parse what is happening here, as ToS is a CwF with extra
structure (def 3.1), but in M, the interpretation of each context is also a CwF
with extra structure (flCwF), so we have "a structured CwF of structured
CwFs". This part is explained and illustrated in detail in the previous
"Constructing QIITs" work, we should direct the reader there more clearly in the
text (sadly we don't have space for similar exposition here)

- "That size issue again"

We can actually eliminate into any type from the (Nat : Set₁) given by the term
model: while the induction principle from Thm 6.10 requires as input a displayed
Nat-algebra at level 1, by cumulativity it is enough to provide a disp. algebra at
level 0. This should work for any particular QIIT. We only
included the general cumulativity proof for plain algebras in def 6.8 and omitted it
for displayed algebras. The displayed case is a bit more fiddly, because we need to handle
cases where the base algebra is at a different level than the displayed algebra itself.
Levels are rather painful, we made an effort to keep the universe setup as simple as
possible, and we made a point to avoid "heterogeneously" leveled constructions such as this.
We'll try to note this in the paper.

- "The category example in Section 1 is not very convincing"

The Cat signature itself requires (finitary) QII features to specify, and it's a
very prominent structure. It happens to be the case that the initial Cat is easy
to construct, but we don't think that detracts from its value as an example; also,
our M semantics provides a bunch of structure unrelated to initial algebras.

- "Smaller comments & typos"

We will try to improve based on these comments

- "Could the usual parametricity arguments be used to show that your
  "Church-encoded" definition of signatures in fact is initial?"

Definitely not internally to cETT, because of the universe polymorphism, but
even without that it's probably not feasible; see "Impredicative encodings of
HITs" by Awodey et al. We don't really need the Church-encoding to be nice and
well-behaved in this paper, we only need a way to write down signatures and
interpret them in a model, similarly to how programmers use "finally tagless"
encodings.

--------------------------------------------------------------------------------

TODO:
- intuitive expl for isofib semantics necessity
- directing readers to CQIITs in sec 6
- improve based on review 3 "Smaller comments"
- response to displayed algebra cumulativity